---
title: "Douwe_egberts_Survey"
author: "Michiel van Eldik"
date: "10/6/2020"
output: pdf_document
---

```{r, message=FALSE }
library(qualtRics)
library(psych)
library(dplyr)
library(epiDisplay)
library(fastDummies)
library(stringr)
```

```{r, message=FALSE }
de.df <- read_survey("DE_1.csv")
```

***

# __"How Frequently do you drink a cup of coffee on average?"__
## __Frequency table__
```{r, message=FALSE }
tab1(de.df$Q27, cum.percent=TRUE, main="Frequency of drinking coffee")
```

***

# __knowledge of the brand__
## __"Have you heard of the brand Douwe Egberts?"__
```{r, message=FALSE}
tab1(de.df$Q6, sort.group="decreasing", cum.percent=TRUE, main="heard of DE")

```

## __"I have an opinion about the brand Douwe Egberts."__
### __Frequency table__
```{r }
tab1(de.df$Q7, cum.percent=TRUE, main="I have an opinion on DE")
```

### __Mean__
```{r }
mean(de.df$Q7)
```
### __Standard Deviation__
```{r }
sd(de.df$Q7)
```


## __"I am well aware of the brand Douwe Egberts"__

### __Frequency table__
```{r }
tab1(de.df$Q8, cum.percent=TRUE, main="I am well aware of DE")
```

### __Mean__
```{r }
mean(de.df$Q8)
```
### __Standard Deviation__
```{r }
sd(de.df$Q8)
```

## __Internal reliability__
```{r }
sc_awareness <- data.frame(scale(de.df$Q8), scale(de.df$Q7))
cor.test(de.df$Q8, de.df$Q7, method="spearman")
```

***
# __Brand Personality Associations Douwe Egberts__
## __Dimension 1, Item 1: Sincerity__
```{r }
tab1(de.df$Q12_1, cum.percent=TRUE, main="Down-to-Earth")
```
```{r }
mean(de.df$Q24)
```
## __Dimension 1, Item 2: Warmth__

```{r }
tab1(de.df$Q12_2, cum.percent=TRUE, main="Warmth")
```

## __Dimension 1 Internal Reliability__
```{r }
sc_d1 <- data.frame(scale(de.df$Q12_1), scale(de.df$Q12_2))
cor.test(de.df$Q12_1, de.df$Q12_2, method="spearman")
```


## __Dimension 2, Item 1: Excitement__
```{r }
tab1(de.df$Q12_3, cum.percent=TRUE, main="Excitement")
```

## __Dimension 2, Item 2: Trendy__
```{r }
tab1(de.df$Q12_4, cum.percent=TRUE, main="Trendy")
```

## __Dimension 2  Internal Reliability__
```{r }
sc_d1 <- data.frame(scale(de.df$Q12_3), scale(de.df$Q12_4))
cor.test(de.df$Q12_3, de.df$Q12_4, method="spearman")
```


## __Dimension 3, Item 1: Competence__
```{r }
tab1(de.df$Q12_5, cum.percent=TRUE, main="Reliable")
```
## __Dimension 3, Item 2: Expertise__
```{r }
tab1(de.df$Q12_6, cum.percent=TRUE, main="Expertise")
```

## __Dimension 3, Internal Reliability__
```{r }
sc_d1 <- data.frame(scale(de.df$Q12_5), scale(de.df$Q12_6))
cor.test(de.df$Q12_5, de.df$Q12_6, method="spearman")
```


## __Dimension 4, Item 1: Sophistication__
```{r }
tab1(de.df$Q12_7, cum.percent=TRUE, main="Upper Class")
```

## __Dimension 4, Item 2: Glamorous__
```{r }
tab1(de.df$Q12_8, cum.percent=TRUE, main="Glamorous")
```

```{r }
sc_d1 <- data.frame(scale(de.df$Q12_7), scale(de.df$Q12_8))
cor.test(de.df$Q12_7, de.df$Q12_8, method="spearman")
```

# __Brand Equity from a Consumers' Perspective__
## __"If there are other brands similar to this one in terms of quality, I would still prefer Douwe Egberts.__
```{r }
tab1(de.df$Q24, cum.percent=TRUE, main="Preference of DE over other, similar brands")
```

### __Mean__
```{r }
mean(de.df$Q24)
```
### __Standard Deviation__
```{r }
sd(de.df$Q24)
```

## __Perceived Likability of the Brand Douwe Egberts__

```{r }
tab1(de.df$Q12_9, cum.percent=TRUE, main="Likeability")
```

### __Mean__
```{r }
mean(de.df$Q12_9)
```
### __Standard Deviation__
```{r }
sd(de.df$Q12_9)
```


## __Perceived Uniqueness of the brand Douwe Egberts__

```{r }
tab1(de.df$Q12_10, cum.percent=TRUE, main="Uniqueness")
```


### __Mean__
```{r }
mean(de.df$Q12_10)
```
### __Standard Deviation__

```{r }
sd(de.df$Q12_10)
```

# Rank 1
```{r }

aaa <- data.frame(as.numeric(strsplit(as.character(de.df$Q39_0_GROUP), "")[[1]]))
aaa <- cbind(aaa, (data.frame(as.numeric(strsplit(as.character(de.df$Q39_0_GROUP), "")[[2]]))))

```



## Summary Pie chart

```{r }
lbls <- c("Sincerity", "Excitement", "Competence", "Sophistication")
listje <- c(
  mean(de.df$Q12_1) +
  mean(de.df$Q12_2),
  mean(de.df$Q12_3) +
  mean(de.df$Q12_4),
  mean(de.df$Q12_5) +
  mean(de.df$Q12_6),
  mean(de.df$Q12_7) +
  mean(de.df$Q12_8)
  )
pie(listje, labels=lbls)
```



```{r }

```

```{r }

```

```{r }

```

