---
title: "Question 2 RMarkdown"
author: "Michiel van Eldik"
date: "9/18/2020"
output: pdf_document
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Setting everything up
Loading packages from libraries
```{r}
library(ggplot2)
library(dplyr)
```

Loading csv file.
```{r}
survey_data <- read.csv("/home/mitchy123/Documents/Statistical learning/SLIMSustainableFashionData.csv", sep= ",", header = TRUE)
sc_data <- survey_data
```

## Getting to know the data
### Boxplot
A multi-facted box plot is a convenient way to get a quick overview of our data.
```{r}
boxplot(WTR ~ Education, data = sc_data,
        xlab = "Education", ylab = "WTR",
        frame = FALSE, col = c("#00AFBB", "#E7B800", "#FC4E07"))
```
To be honest, I don't get much wiser from looking at this. Apart from the fact that it suggests that "5" is not a very common score.

## Table of means, counts, standard deviations per educational level for WTR variable
```{r}
new_data <- sc_data %>%
  group_by(Education) %>%
  summarise(
    count = n(),
    mean = mean(WTR),
    sd = sd(WTR))
new_data
```
We can conclude from this that the counts per educational level vary with relatively little counts for educational levels 1 and 2. 
Furthermore, means and standard deviations don't seem to vary significantly. 
This might foreshadow that Educational level is not a significant factor in predicting willingness to recommend.

## Homogeneity of variances tests 
An assumption for ANOVA is homogeneity of variances. This can be done in different ways. I chose Bartlett's and the more robust Fligner's test. Levene's test is also an option but I didn't manage to install the package for it correctly as of yet.

### Bartlett's test
```{r}
res <- bartlett.test(WTR ~ Education, data = sc_data)
res
```

### Fligner's test
```{r}
flig <- fligner.test(WTR ~ Education, data = sc_data)
flig
```
Both tests' null hypotheses are homogeneous variances. In either case, we can't reject this null because p-values are too high.
This means that the variance of variances across educational levels is insignificant.

## Normality
Another assumption for ANOVA is normality. However, our dependent variable WTR is a discrete variable with only 5 levels that we treat as continuous. Nonetheless, we can take a look at the qq plots.

```{r}
# Function for plotting qq norms
WTR_edu_qq <- function(sc_data, edu_level) {
  newie <- sc_data %>%
    filter(Education == edu_level)
  qqnorm(newie$WTR)
}

# Using the qq norm function by iterating over all education levels
for (value in c(1, 2, 3, 4)) {
WTR_edu_qq(sc_data, value)  
}
```

## ANOVA one-way

```{r}
WTR_AOV <- aov(WTR ~ Education, data = sc_data)
summary(WTR_AOV)
```

## ANCOVA applied using linear model

```{r}
model_1 = lm (WTR ~ Education + PurchaseFreq,
            data = sc_data)
summary(model_1)
```


